---
title: "Exploratory Data Analysis"
output: 
  html_notebook:
    fig_width: 6 
    fig_height: 4
    
---


```{r}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```


# Load Libraries

```{r}
library(tidyverse)
library(descr)
library(Amelia)
library(lubridate)
library(ggmap)
library(quantreg)
library(shapefiles)
library(leaflet)
library(stringi)
library(ngram)
library(qdap)
library(rgdal)
library(plotly)
library(data.table)
```

# Load Data

```{r}
business <- read_csv("yelp_toronto_business.csv")
review <- read_csv("yelp_toronto_review.csv")
```

# Exploratory Data Analysis

Our focus will be on the business data, but we will use the review data to get some insights about businesses and will do some feature engineering to create useful variables for the model.

## review data

```{r}
glimpse(review)
```

```{r}
# Are there any missing values in the review data?
paste("There is :", sum(is.na(review)), "missing values in the review data.")
```
Sounds good, there are no missing values in the review dataset. Later we will check if some variable will help for building our model.
For that data set, we will select later 4 features ('business_id', 'stars', 'text', 'date') and try to create new features and add them to the initial business data. 

```{r}
#Create a new variable containing the number of word and characters in reviews 
review <- review %>%
            mutate(no_characters = nchar(review$text), no_words = str_count(review$text, "\\w+")) #%>%
            #group_by(business_id, word_count) %>%
            #arrange(desc(word_count))
                  
```


## Business Data

Now let check the business data set and respond to the questions below:

Does the data contain some missing values?
What is the ditribution of the variables?
How are the different variables related to each other?

```{r}
# Business data structure
glimpse(business)
```

### Visualization of missing values

```{r}
# let visualize the percentage of missing values for each feature

missing <- data.table(pmiss = sapply(business, function(x) { (sum(is.na(x)) / length(x)) }),
                      column = names(business))
  
p <- ggplot(missing,aes(x = reorder(column, -pmiss), y = pmiss)) +
      geom_bar(stat = 'identity', fill = 'steelblue') + 
      scale_y_continuous(labels = scales::percent) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
      labs(
        x = 'Feature',
        y = '% missing',
        title = "Missing data by feature"
        )


ggplotly(p) 
```

We have that summary for the business data:
18,233 Observations and 13 Variables  
$ address      character class with 14422 unique values and 283 missing values  
$ business_id  character class with 18233 unique values and 0 missing values  
$ categories   character class with 10028 unique values and 33 missing values  
$ city         character class with 1 unique values and 0 missing values  
$ is_open      integer class with 2 unique values and 0 missing values  
$ latitude     numeric class with 15366 unique values and 1 missing values  
$ longitude    numeric class with 15315 unique values and 1 missing values  
$ name         character class with 15292 unique values and 0 missing values  
$ neighborhood character class with 80 unique values and 3435 missing values  
$ postal_code  character class with 5261 unique values and 117 missing values  
$ review_count integer class with 380 unique values and 0 missing values  
$ stars        numeric class with 9 unique values and 0 missing values  
$ state        character class with 2 unique values and 0 missing values  


The neighborhood variable have the most missing values (19%). 

```{r}
# correcting coordinates
business[which(business$address == "2138 Queen Street E"),] <- business %>% 
                                                        filter(business$address == "2138 Queen Street E") %>%
                                                        mutate(longitude = -79.293425, latitude = 43.671584)

#adding coordinates for the postal code M5H 4G1 (lon: -79.3854, lat: 43.6508)  
business[which(is.na(business$longitude) == TRUE),] <- business %>% 
                                                        filter(is.na(longitude) == TRUE) %>%
                                                        mutate(longitude = -79.3854, latitude = 43.6508)

# converting "is_open" attribute from integer to factor and change values to yes or no
#lowering some character variable, converting "stars" from int to factor

business <- business %>% 
              mutate(is_open = as_factor(if_else(as.logical(is_open) == TRUE, "yes", "no")), 
                     categories = str_to_lower(categories),
                     name = str_trim(str_to_lower(name),side = 'both'),
                     neighborhood = str_to_lower(neighborhood),
                     stars =ordered(as_factor(as.character(stars)), levels = c("1", "1.5", "2", "2.5", "3", "3.5", "4", "4.5", "5"))
                     ) 

```

### Univariate and bivariate analysis

Let check the distribution in the target variable. How many businesses are open? How many are closed?

```{r}
# Target variable distribution
p <- ggplot(business, aes(x = is_open)) +
      geom_bar(aes(fill = is_open))

ggplotly(p)
```

14023 businesses are open (77%) and 4210 closed (23%). Our target variable is imbalanced.

```{r}
# Star rating distribution
p <- ggplot(business, mapping = aes(stars)) +
      geom_bar(aes(fill = stars)) +
      labs(title ="Star rating distribution")

ggplotly(p)
```
Around 70% of the businesses got a rating between 3 and 4.5








```{r}
# The distribution of number of reviews
#we have applied log since the distribution is skewed
p <- ggplot(data = business, aes(x = review_count)) + 
      geom_histogram(bins = 50, fill = "darkgreen", binwidth = 25) +
      #geom_density(alpha=.2, fill="#FF6666") +
      scale_y_log10() +
      labs(title ="Distribution of reviews")


ggplotly(p)
#ggsave("reviews_dist.png", width = 10, height = 5)

```



Let visualize if "closed" Businesses received less reviews, and if the closure depend on rating. 

```{r}
# visualizing the number of review per rating star

p1 <- ggplotly(
       business %>%
          select(stars, review_count) %>%
          group_by(stars) %>%
          summarize(review_count = sum(review_count)) %>%
        ggplot(stars, mapping = aes(stars, review_count)) +
          geom_col(fill = 'darkgreen') +
          labs( x = "Star rating", y = "Number of reviews", title ="Number of reviews per rating star")
        )

p2 <- ggplotly(
       business %>%
          select(stars, is_open, review_count) %>%
          group_by(stars, is_open) %>%
          summarize(review_count = sum(review_count))%>%
        ggplot(business, mapping = aes(stars, review_count, fill = is_open)) +
          geom_col(position = 'dodge') +
          labs( x = "Star rating", y = "Number of reviews", title ="Number of reviews per rating star")
        )


subplot(p1, p2, nrows = 2, margin = 0.02, shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE)

```

From the graph above, we can gather that the distribution of number of reviews is quite the same for 'open' businesses and 'closed'  businesses. As the target variable is not balanced, it's not a supprise that for each rating we observed a significant difference in term of number of reviews.


```{r}
# visualizing the number of business per rating star

p1 <- ggplotly(
        ggplot(business, mapping = aes(stars)) +
        geom_bar(fill = "darkblue") +
        labs( x = "Star rating", y = "Number of businesses", title ="Star rating distribution business open/closed")
        )



p2 <- ggplotly(
        ggplot(business, mapping = aes(stars, fill = is_open)) +
        geom_bar(position = "dodge") +
        labs( x = "Star rating", y = "Number of businesses", title ="Star rating distribution business open/closed")
        )


subplot(p1, p2, nrows = 2, margin = 0.02, shareX = TRUE, shareY = TRUE, titleX = TRUE, titleY = TRUE)
#ggsave("stars_dist2.png", width = 5, height = 5)
```

One assumption we can make is that ratings has no impact on a business closure. 


### Geographical visualization

```{r}
center_long = median(business$longitude, na.rm = TRUE)
center_lat = median(business$latitude, na.rm = TRUE)

leaflet(business) %>% 
      addTiles()%>%
      #addProviderTiles("Esri.NatGeoWorldMap") %>%
      addCircles(lng = ~ longitude, lat = ~latitude, radius = ~sqrt(review_count)) %>%
      setView(lng = center_long, lat = center_lat, zoom = 10)

```

In the Great Toronto Area (GTA), most of the businesses are located in Downtown Toronto.


### Leading Business Categories:

Let's look at the leading business categories in our data set. A business is linked to multiple categories in our dataset, so we have to do a bit of preprocessing, which is simple using the dplyr package.

```{r}
# Top categories of business in GTA

categorie <- business %>% 
                unnest(categories = str_split(categories, ",")) %>%
                mutate(categories = str_trim(categories,side = "both")) %>%
                select(categories) %>% 
                group_by(categories) %>% 
                summarise(n=n()) %>% 
                arrange(desc(n)) %>% 
                head(25)

p <- ggplot(categorie, aes(x = reorder(categories, n), y = n)) +
      geom_col(aes(fill = n)) +
      scale_fill_gradientn(colours=RColorBrewer::brewer.pal(11,"Spectral")) +
      coord_flip() +
      labs(title ="Top 25 business categories")



ggplotly(p)
```
More than 60% of the rated businesses rated have their activities related to food.

__________________

As in teh data set each business is linked to multiple categories, we will create a new feature to count the number of categories per business.

```{r}
# Create a new variable to gather the number of categories for each business
business <-  business %>% 
              mutate(cat_count = length(business$categories))
                            



```



# Top Business names


```{r}
business_name <- business %>% 
                select(name) %>% 
                group_by(name) %>% 
                summarise(n=n()) %>% 
                arrange(desc(n)) %>% 
                head(25)

p <- ggplot(business_name, aes(x = reorder(name, n), y = n)) +
      geom_col(aes(fill = n)) +
      scale_fill_gradientn(colours=RColorBrewer::brewer.pal(11,"Spectral")) +
      coord_flip() +
      labs(title ="Top 25 business names")



ggplotly(p)
```

```{r}

```



