---
title: "Reviews Analysis"
output:
  html_notebook:
    fig_height: 8
    fig_width: 12
  html_document:
    df_print: paged
---

In this notebook, I explore the review data set and do some feature engineering for the business data which will be used to train my model.

```{r}
library(tidyverse)
library(descr)
library(lubridate)
library(stringi)
library(ngram)
library(rgdal)
library(plotly)
library(data.table)
library(corrplot)
```

# Loading review data

```{r}
review <- read_csv("yelp_reviews_sentiment.csv") # from the IPython NoteBook 'Yelp_reviews_sentiment.ipynb' where I created a feature to get the sentiment score of each review. I used the ntlk library to do it.
```

```{r}
glimpse(review) #visualizing the structure of the review data
```

```{r}
# Are there any missing values in the review data?
paste("There is :", sum(is.na(review)), "missing values in the review data.")
```
_____________

Sounds good, there are no missing values in the review dataset. Later I will check if some variable will help for building our model. For that, I selected the following variables in the review data: 'business_id', 'stars', 'text', 'useful', 'funny', 'cool', 'senti_score'. And I tried to create new features to add to the initial business data.

```{r}
review <- review %>%
            mutate(num_char = str_length(text), num_words = sapply((str_split(text, boundary ('word'))), length), 
                   senti_score = round(senti_score, digits = 3)) %>%
            select(-c('review_id', 'user_id' , 'date', 'text'))

glimpse(review)

corrplot(cor(review[-1]), method = 'number')
```



Obviously, number of characters and number of words in a review are correlated. From the graph above, we observe a positive correlation between sentiment score and the rating gave by reviewers. That means a postive sentiment is likely to lead to a higher rating star. And there is a negative weak correlation between sentiment score and number of words/character. We can interpret that as people who had a positive experience with a business tended to write short review.  
The variables 'useful', 'funny', and 'cool' represent the number of votes received qualifying a review as respectively useful, funny or cool. 

Now, let visualize the relation between sentiment score and number of words

```{r}
ggplot(review , aes( senti_score)) +
  geom_histogram(fill='darkgreen', color = 'black', binwidth = .02)
```




```{r}
#plot_ly(data = review %>% filter(num_words > 0), x = ~num_words, y = ~senti_score)


ggplot(review %>% filter(num_words > 0) , aes(num_words, senti_score)) +
  geom_point(color='darkblue', position = 'jitter') + 
  scale_x_log10() +
  labs(
    x = 'number of words per review',
    y = 'sentiment score per review',
    title = "Relation between sentiment score and number of words"
     )
```

We observe that as the number of word increased the sentiment decreased from the positive side to zero and increased from the negative side to zero. We can gather that when a review is long the sentiment is not clear to gather.
Let visualize the relation between number of words and 'useful'/'funny'/'cool'.

```{r}
ggplot(review %>% filter(num_words > 0) , aes(num_words, useful)) +
  geom_point(color='darkblue', position = 'jitter') + 
  scale_x_log10()

ggplot(review %>% filter(num_words > 0) , aes(num_words, funny)) +
  geom_point(color='darkblue', position = 'jitter') + 
  scale_x_log10()

ggplot(review %>% filter(num_words > 0) , aes(num_words, cool)) +
  geom_point(color='darkblue', position = 'jitter') + 
  scale_x_log10()
```

```{r}
head(review)
```
____________
As I want to make my prediction on businesses, I will group the reviews by businesses and aggregate the different variables.
I decided to selected only the senti_score and number of words to add to the business data.

```{r}
# Grouping reviews by businesses 
sentiment <- review %>%
            group_by(business_id) %>% 
            summarise(
                      avg_words = round(mean(num_words), 0),
                      avg_sent = round(mean(senti_score), 3),
                      )
write_csv(sentiment, "sentiment.csv")
                  
```

