---
title: "Filter to Toronto reviews"
output: html_notebook
---

We will subset the different datasets to get observations related to Toronto businesses.

# Libraries

```{r}
library(tidyverse)
```

# Loading the yelp datasets

```{r}
business <- read_rds("business_data.rds")
review <- read_rds("review_data.rds")
user <- read_rds("user_data.rds")
checkin <- read_rds("checkin_data.rds")
tip <- read_rds("tip_data.rds")
photo <- read_rds("photo_data.rds")
```

# Filter to businesses in Toronto

I decided to create 2 data for Toronton businesses. One with the features "attribute" and "hours" (will remain in R object for later use) and another without those features (to be converted in csv file). 

```{r}
toronto_business <- business[which(business$city == "Toronto"),]

toronto_business_2 <- business %>%
                        select(-attributes, -hours) %>%
                        filter(city == "Toronto")
```


# Filter review on Toronto Businesses

I will do a semi-join on the column "business_id" between review and business to get reviews done on toronto businesses. And so one for the other datasets except checkin.

```{r}
toronto_review <- semi_join(review, toronto_business, by = "business_id")
```

# Toronto tips

```{r}
toronto_tip <- semi_join(tip, toronto_business, by = "business_id")
```

# Toronto reviewers : User who made reviews on Toronto Businesses

```{r}
toronto_user <- user %>%
                  semi_join(toronto_review, by = "user_id")
```

# Toronto business photo

```{r}
toronto_photo <- semi_join(photo, toronto_business, by = "business_id")
```

# Saving in csv files

```{r}
write_csv(toronto_business_2, "yelp_toronto_business.csv")
write_csv(toronto_photo, "yelp_toronto_photo.csv")
write_csv(toronto_review, "yelp_toronto_review.csv")
write_csv(toronto_tip, "yelp_toronto_tip.csv")
write_csv(toronto_user, "yelp_toronto_user.csv")
```

# Saving in R object

```{r}
#saveRDS(toronto_business_2, "yelp_toronto_business.rds")
#saveRDS(toronto_photo, "yelp_toronto_photo.rds")
#saveRDS(toronto_review, "yelp_toronto_review.rds")
#saveRDS(toronto_tip, "yelp_toronto_tip.rds")
#saveRDS(toronto_user, "yelp_toronto_user.rds")
```

